Dynamics of the Lower Solar Atmosphere

### 2024-07-22 chapter5/ms.tex
Updates from Overleaf

%UNLIKE IN A REGULAR TEX FILE, DON'T PUT ANY PREAMBLE MATERIAL HERE

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Aditya-L1}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Solar Ultraviolet Imaging Telescope}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{SUIT \& Solar Flares}\label{sec:suit_and_flare}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Solar flares are the most powerful magnetic events in the solar system. They are described as a sudden increase in brightness in localized areas on Sun. Within tens of minutes, they can release over $10^{32}$ erg of energy, which is emitted across the entire electromagnetic spectrum from radio to gamma rays. They can also launch high energetic particles into the interplanetary medium. Most of the flares occur in magnetic active regions, and the amount of flare energy released is comparable to the free energy stored in the magnetic system. The term "flare" is generally used explicitly for the entire magnetically-driven event's electromagnetic radiation, as it is the most significant fraction of the total energy liberated. The total energy released varies from event to event. It is also known that larger events occur much less frequently than smaller events.

The Solar Ultraviolet Imaging Telescope \citep[SUIT;][]{ghosh16,article} is one of the seven payloads onboard the Aditya-L1 mission \citep{adityal1} of the Indian Space Research Organization (ISRO). With its 11 science filters (3 broadband and eight narrowband), SUIT will have the capability to probe different heights of solar atmosphere in photosphere and chromosphere to help us understand the various processes that transport mass and energy from one layer to another. SUIT will provide full disk as well as partial disk images of the Sun with a pixel size of 0.7". Through SUIT imaging, we would be able to resolve solar flares spatially on the surface of the Sun, for the first time in near ultra-violet (NUV), which will help us to address the questions regarding their build-up and triggering mechanism. In addition, to measure the spatially resolved solar spectral irradiance within the wavelength range that is central for studying the Chemistry of oxygen and ozone in the Stratosphere of Earth's atmosphere.

It has been shown that the majority of flare energy emerges at the visible and UV wavelength range \citep{woods06}. \cite{woods04} showed that about 77\% of the energy is released in the wavelength range > 200 nm, and only ~ 23\% is seen in extreme ultraviolet (EUV) and soft X-ray (SXR), i.e., below 200 nm \citep{Nei_1989,neidig93,kretzschmar11}. Although the energy content in hard X-ray (HXR) is a tiny fraction of the total energy budget, they are still crucial in understanding the energization process \citep{holeman11}. However, to develop a comprehensive understanding of solar flares, it is mandatory to perform multi-wavelength studies of all kinds of flares. This may have implications on the physics of the origin of solar flares and different physics processes and contribute to the solar spectral irradiance as a function of the solar activity cycle. Although we have been observing Sun and Solar flares in various wavelengths, the spectral energy distribution of the radiated energy from the flares is still very poorly understood. The first solar flares were observed from the ground in the visible domain \citep{carrington1859,neidig93}. It is also well known that the flare emission in the visible domain occurs mainly in $H\alpha$ and Ca \Romannum{2} lines \citep{canfield90,falchi92,heinzel94}. However, the lesser understood component of the visible and Near Ultra Violet (NUV) emission is the enhancement of the continuum. The study of the white-light (WL) flares has proven to be very difficult because they have a very short duration and low contrast against the background, making their observation from Earth rare and of poor quality. Also, the flares in NUV are not observable from any ground-based instrumentation as most of the NUV gets absorbed in the upper atmosphere, thus requires space-based observations.

The origin of the WLFs, i.e., the physical process responsible for generating the continuum and its contribution to the overall energy distribution, is still highly uncertain. The question remains whether WLFs are photospheric phenomena due to $H^{-}$ free-free emission or chromospheric phenomena due to H free-bound emission. The more recent studies further constrain the origin of the WLFs to be Chromospheric phenomena, as it has been shown that the WL and HXR footpoint centroids are cospatial. Similarly, \cite{krucker15} constrained the cospatial WL and HXR footpoints within the chromosphere for three flares. With the help of SUIT, we would be able to localize the WL flares and resolve them on various parts of the Solar disk, which along with observations from Interface Region Imaging Telescope (IRIS), Helioseismic and Magnetic Imager (HMI), would help us localize the source of the WL footpoints and also comment on the formation mechanism of the WL itself.

\begin{figure}
    \centering
    \includegraphics[width = \linewidth]{Figures/Krucker_2015_ApJ_802_19_page-0004.jpg}
    \caption{X-ray and optical imaging of the three flares at the peak time of the impulsive phase: the images are from HMI with the pre-flare image subtracted. The contours represent RHESSI Clean maps in the thermal (red, 12–15 keV) and non-thermal (blue, 30–80 keV) HXR range \citep{krucker15}.}
    \label{fig1}
\end{figure}

The spectral energy distribution of flares is one of the critical areas of interest in the physics of flares. A complete understanding of this will help us decode the physical processes involved in solar flares and help quantify their effects on solar spectral irradiance. Ideally, it would be essential to observe flares at all wavelengths simultaneously with sufficient spatiotemporal resolution to figure out the spectral energy distribution of flares. Unfortunately, this is generally not the case, and we have to rely on the sporadic observations made using ground-based instruments in the visible domain. As mentioned earlier, the majority of the flare energy is emitted in the NUV and visible domain. \cite{kretzschmar10,kretzschmar11}, performed statistical studies using a large number of flare observations across a wide energy range. They demonstrated, at the peak of the flare, about 70\% of the total energy was radiated in the continuum visible and NUV channel as illustrated in \ref{tab1}. SUIT would observe and resolve the solar flares within 200-400 nm using 8 NBs and 3 BBs. This, along with IRIS data, would help us comment on the energy distribution of flares of various classes.

%%--------------------------------------------%%
\begin{table}[]
    \centering
    \resizebox{\textwidth}{!}{%
    \begin{tabular}{||m{.1\linewidth}|m{.15\linewidth}|m{.17\linewidth}|m{.17\linewidth}|m{.17\linewidth}|m{.15\linewidth}||}
    \hline
    \hline
        Mean X-ray class & TSI(ergs) & Ratio $\frac{26-34nm}{TSI}$ & Ratio $\frac{0-50nm}{TSI}$ & Ratio $\frac{0.1-0.8nm}{TSI}$ & Ratio $\frac{continuum}{TSI}$\\
        \hline
        X3.2 & 5.9$\times~10^{31}$ & 0.9-0.8\% & 12-9\% & 1.2-1\% & 67\% \\
        M9.1 & 1.6$\times~10^{31}$ & 1.7-0.4\% & 23-5\% & 1-0.4\% & 85\% \\
        M4.2 & 1.3$\times~10^{31}$ & 2.2-0.5\% & 18-6\% & 0.6-0.3\% & 74\% \\
        M2.0 & 5.1$\times~10^{30}$ & 1.7-0.6\% & 18-6\% & 0.7-0.4\% & 69\% \\
        C8.7 & 3.6$\times~10^{30}$ & 1.5-0.5\% & 16-5\% & 0.4-0.2\% & 72\% \\
        \hline
    \end{tabular}}
    \caption{Spectral Energy Distribution from a sample of 2100 flares across various wavelengths\citep{kretzschmar11}.}
    \label{tab1}
\end{table}
%%--------------------------------------------%%

Finally, one of the major question of interest is how the Solar flares affect the Spectral Solar Irradiance (SSI) and Total Solar Irradiance (TSI) variability from a short to much longer, Solar Cycle timescale. \cite{kretzschmar10,kretzschmar11} performed statistical studies using a large number of flare observations across a wide energy range. For this purpose, they used the full Sun observations of Solar flux from Solar and Heliospheric Observatory (SoHO), three visible Solar irradiances from VIRGO/Solar Photometer (SPM) passbands centred on 402 nm, 500 nm, and 862 nm, respectively, from 1996 to 2008. Additionally, they also use the EUV irradiance in the ranges 0.1-50 nm and 26-34 nm measured by SOHO/Solar EUV Monitor \citep{judge98} and SXR measurements from GOES satellites. They showed that stacked TSI variation profiles during solar flares show variation at more than 2 sigma level during the peak flare time, indicating the presence of flare signals in the TSI measurements.

\begin{figure}[h!]
    \centering
    \includegraphics[width = \linewidth]{Figures/nphys1741_page-0002.jpg}
    \caption{Averaged TSI variations during flares. The TSI time-series averages over four exclusive sets of solar flares of decreasing amplitude. The black and pink curves correspond respectively to the TSI measured by the PMOV6 and the DIARAD radiometers. The dashed lines correspond to the 95\% confidence levels, while the vertical line denotes the peak time of the flare \citep{kretzschmar10}.}
    \label{fig2}
\end{figure}

This allows us to quantify the solar variability induced be solar flares of various timescales. With the help of SUIT, we would be able to localize the flare locations and study the change in SSI from the local environment in the 11 science filters. This information combined with TSI measurements can allow us to quantify the effect of flares of various energy scales on the TSI variability of the Sun. As, both the TSI and SSI variability directly or indirectly couples with various atmospheric parameters, we can also study the effect the flares have on them. For example, the Earth's atmospheric chemistry and composition respond to any changes in solar UV output in a very nonlinear fashion \citep{haigh07}. Since the number of flares also shows a change with solar activity, it is prudent to ask how much flares contribute to the solar spectral irradiance in NUV. This is particularly important because the irradiance in NUV plays a key role in heating the upper and middle layers of the Earth's atmosphere directly and their coupling with Stratosphere. It also directly influences the middle and lower atmosphere chemistry and composition via the Ozone-Oxygen cycle.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Throughput model and imaging performance of SUIT}\label{sec:suit_imaging}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Filter Choice for SUIT}\label{sec:filter_choice}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Photometric, Radiometric \& Spectral Calibration of SUIT}\label{sec:suit_cal}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%NOTE THE LACK OF A BIBLIOGRAPHY CALL IN THIS FILE. BIBLIOGRAPHY WORK HAPPENS OUTSIDE THE CHAPTER TEX FILES.

>>>>

%\begin{abstract}

 %   The  \ion{Mg}{2}~k \& h line intensity ratios can be used to probe the characteristics of the plasma in the solar atmosphere. In this study, using the observations recorded by the Interface Region Imaging Spectrometer (IRIS), we study the variation of the  \ion{Mg}{2}~k \& h intensity ratio for three flares belonging to X-class, M-class, and C-class, throughout their evolution. We also study the k-to-h intensity ratio as a function of magnetic flux density obtained from the line-of-sight magnetograms recorded by the Helioseismic and Magnetic Imager (HMI) on board the Solar Dynamics Observatory (SDO). Our results reveal that while the intensity ratios are independent of magnetic flux density, they show significant changes during the evolution of the C-class and M-class flares. The intensity ratios start to increase at the start of the flare and peak during the impulsive phase before the flare peak and decrease rapidly thereafter. The values of the ratios fall even below the pre-flare level during the peak and decline phases of the flare. These results are important in the light of heating and cooling of localized plasma and provide further constraint on the understanding of flare physics.

%\end{abstract}
\justifying

%%----------------------------------------------------
\section{Introduction} \label{sec:intro}
%%----------------------------------------------------

Solar flares are the most energetic events on the Sun, where an enormous amount of magnetic free energy is released due to the reconfiguration of the coronal magnetic field. The released energy can cause particle acceleration, heating and flows in the solar atmosphere and a transient enhancement in solar radiative output. Notably, a significant portion of the radiated energy during flares originates from the dense chromosphere \citep{fletcher10,milligan14}. Therefore, examining chromospheric lines during flares offers valuable diagnostic tools for understanding the physics of solar flares and their impact on the local plasma environment.

The chromosphere emits radiation across various ultraviolet (UV) and optical lines. While many optical lines, such as H$\alpha$ and \ion{Ca}{2}, are routinely observed from ground-based telescopes, observations of the  \ion{Mg}{2} resonance lines have been relatively infrequent in the past. However, since the launch of the Interface Region Imaging Spectrograph (IRIS) \citep{iris}, regular monitoring of these lines with excellent spatial and spectral resolution has become possible.

The  \ion{Mg}{2} k and h lines represent transitions to the ground state from finely split upper levels ($3p~^{2}P_{\nicefrac{3}{2}}${--}$3s~^{2}S_{\nicefrac{1}{2}}$ and $3p ^{2}P_{\nicefrac{1}{2}}${--}$3s^{2}S_{\nicefrac{1}{2}}$), resulting in optically thick lines at wavelengths 2796.34~{\AA} ( \ion{Mg}{2} k) and 2803.52{\AA} ( \ion{Mg}{2} h). It has been suggested that the intensity ratios of these lines can offer insights into the optical depth of the local environment \citep{kerr15}.

The integrated intensity of a line transitioning from an upper level $j$ to a lower level $i$ depends on the collision strength $\Omega_{ij}$ for that transition, given by~\citep{henri62,mariska92},

%%---------------------------------------------------------------
\begin{equation*}
\Omega_{ij}=\frac{8\pi}{\sqrt{3}}~\frac{I_{H}}{\Delta \epsilon_{ij}}g\omega_{i}~f_{ij}
\end{equation*}
%%---------------------------------------------------------------

\noindent Here, $I_{H}$ denotes the ionization energy of hydrogen, $\Delta \epsilon_{ij}$ represents the threshold energy for the transition, $g$ is the Gaunt factor, $\omega_{i}$ is the statistical weight of the level, and $f_{ij}$ stands for the oscillator strength. In optically thin conditions, the intensity ratio of the k to h line equals the ratio of collision strengths, as the escape probability of photons is unity. As the  \ion{Mg}{2} k and h lines share the same ionization state and originate from a transition to a shared lower level, and given that the statistical weight ($\omega_{i}$) is the same in both cases, the line intensity ratio is simply the ratio of oscillator strengths ($f_{ij}$). Consequently, this ratio is expected to be 2:1 in optically thin conditions and lower when the medium is optically thick \citep{kerr15,levens19}.

Moreover, the  \ion{Mg}{2} k and h lines can serve to estimate velocity in the middle and upper chromosphere, chromospheric velocity gradients, and temperature in the middle chromosphere \citep{leenarts13a,leenarts13b,pereira13}. Emission from the  \ion{Mg}{2} triplets can help identify heating in the lower chromosphere \citep{pereira15}. Various studies have demonstrated spatial variations in  \ion{Mg}{2} line profiles \citep{dalda23,panos18}. For instance, \cite{polito23} associated the leading edge of flare ribbons with enhanced broadening and strong central reversal, interpreting this difference in profile as indicative of distinct heating mechanisms at different locations within flare ribbons. Similarly, \cite{panos21,panos21_2} revealed differences in line profiles and energy input.

Using observations recorded by the OSO-8 LPSP instrument, \citep{lemaire84} investigated the evolution of intensity ratios of  \ion{Mg}{2} h \& k, \ion{Ca}{2} h \& k, and Ly$\alpha$ \&~$\beta$ lines. They observed that the intensity ratio of the \ion{Ca}{2}~k/h lines increased from 1 to 1.2 during the ascending phase of a flare and returned to 1 during later phases. This correlated temporal behavior across various elements was interpreted as an indication of downward energy propagation, suggesting a potential decrease in opacity due to localized heating at the formation height of the \ion{Ca}{2} line during the flare's rise phase.

Here, we investigate the evolution of intensity ratios of the  \ion{Mg}{2} h \& k lines during three flares: C-class, M-class, and X-class. Specifically, we focus on the dependence of line ratios on the underlying magnetic field strength, a relationship that, to our knowledge, has not been explored previously. The remainder of this chapter is organized as follows. Section \ref{sec:obs} presents the observations utilized in this study, followed by our data reduction and analysis methods, and the results in Section \ref{sec:dar}.

%%----------------------------------------------------
\section{Observations} \label{sec:obs}
%%----------------------------------------------------
%%-------------------------------------------------------
\begin{table*}[ht!]
\centering
\begin{tabular}{|c|c|c|c|c|c|}
\hline
Event & Flare & Flare & Raster & Raster Step & Raster \\
Date & Peak (UT) & Location (arcsec) & Details & (arcsec) & Cadence (s)\\
\hline
Nov 4, 2015  & 13:52 & [37",61"] & Coarse & 2" & 50\\
(M-class) & & & 16-step  & & \\
 & & & & & \\
Oct 22, 2014 & 14:28 & [-292",-302"] & Coarse & 2" & 131\\
(X-class) & & & 8-step  & & \\
 & & & & & \\
Feb 3, 2015 & 22:55 & [198",213"] & Dense & 0.35" & 33\\
(C-class) & & & 16-step  & & \\
\hline
\end{tabular}%}
%\end{center}
\caption{List of flares studied in this chapter.}
\label{tab:my_label}
\end{table*}
%%-------------------------------------------------------

For this study, we selected three flares of M, X, and C classes, as listed in Table~\ref{tab:my_label} by IRIS. IRIS is a NASA small explorer-class solar observation satellite that obtains UV spectra with high spatial (0.33{--}0.4\arcsec per pixel), temporal (1s), and spectral resolution ($\sim$26 and $\sim$53~m{\AA}). The primary lines regularly observed by IRIS include \ion{C}{2},  \ion{Mg}{2}, and \ion{Si}{4}. In the imaging channel, it typically observes in  \ion{Mg}{2} and \ion{Si}{4}. In our study, we utilized observations recorded in the  \ion{Mg}{2} h\& k lines.

As mentioned earlier, this study aims to analyze the evolution of intensity ratios concerning magnetic flux density during the flares of various classes. To achieve this, we incorporated line-of-sight (LOS) magnetic field measurements from the Helioseismic and Magnetic Imager \citep[HMI;][]{hmi} on the Solar Dynamics Observatory \citep[SDO;][]{sdo}. We utilized observations taken at 1600~{\AA} by the Atmospheric Imaging Assembly \citep[AIA;][]{aia}, also onboard SDO, to co-align the IRIS observations with those from AIA and subsequently HMI.

%%----------------------------------------------------
\section{Data analysis and results} \label{sec:dar}
%%----------------------------------------------------
\subsection{M3.7 Flare Observed on Nov 4, 2015}
%%%----------------------------------------------
%%--------------------------------------------------

NOAA AR 12443 generated a multi-ribbon GOES class M3.7 flare on November 4, 2015, which commenced around 13:31 UT and peaked at approximately 13:52 UT, as observed from the GOES Soft X-ray (SXR) 1{--}8~{\AA} flux (Fig.\ref{flare1}a). This event occurred at approximately [37\arcsec,61\arcsec] heliographic position and was extensively observed by IRIS, AIA, and HMI. Fig.\ref{flare1}a illustrates the GOES flux plot of the flare in the 0.5{--}4~{\AA} range (blue) and the 1.0{--}8.0~{\AA} range (red). Fig.\ref{flare1} b depicts AIA 1600{\AA} image of the flaring region, showing two ribbons indicated by arrows. In panel (c), we present the line-of-sight (LOS) magnetic flux density map obtained from HMI, recorded nearly simultaneously with the AIA image shown in panel (b). The white (black) box overlaid on Fig.\ref{flare1}(b) (c) represents the IRIS SJI (Slit-Jaw Imager) field of view (FOV). The white dot-dashed (magenta dashed) box in Fig.\ref{flare1}b (c) indicates the IRIS raster FOV. The FOV of the SJI (approximately [120\arcsec~$\times$119\arcsec]) covers the central part of the flaring region, with a spectral sampling of approximately 0.05~{\AA}/pixel. \cite{li17} investigated the dynamics of the ribbons for this flare, while \cite{karlick18} studied the associated radio bursts.

%%--------------------------------------------------
\begin{figure*}[ht!]
    \centering
\includegraphics[trim={7cm 5cm 7.7cm 6cm},clip,width=\textwidth]{Figures/Flare_M_Nov04_2015_2.eps}
\caption{The M3.7 flare observed on November 4th, 2015. Panel a: GOES flux plot in 0.5{--}4~{\AA} (blue) and 1.0{--}8.0~{\AA} (red). Panel b: AIA 1600~{\AA} image of the flaring region. Arrows locate the primary ribbons. Panel c: LOS magnetic flux density map obtained from HMI near the peak of the flare. The over-plotted white (black) boxes in panel b (c) represent the IRIS SJI FOV. The over-plotted white dot-dashed (magenta dashed) box in panel b (c) shows the IRIS raster FOV.}\label{flare1}
\end{figure*}
%%--------------------------------------------------

Figure~\ref{flare_m_ev} illustrates the evolution of the flare in AIA 304~{\AA}. The flare is connected with a pre-existing filament that undergoes an eruption, splitting into two structures denoted as F1 \& F2 in Fig.\ref{flare_m_ev}(b) \& (c). These two filament structures diverge from each other in opposite directions. The flare generates two primary flare ribbons, labeled as R1 \& R2 in Fig.\ref{flare_m_ev}(b), (c) \& (d). Starting around 13:32 UT, R1 travels southeastward, crossing the IRIS raster FOV, indicated by the white dotted box in Fig.\ref{flare_m_ev}(a), (b), (c) \& (d). The IRIS raster monitors the movement of the northern ribbon R1 and the eastern edge of R2. An animated version of Fig.\ref{flare_m_ev} is available in the online journal for further details.

%%%--------------
\begin{figure*}[ht!]
    \begin{center}
    \includegraphics[width=\textwidth]{Figures/nov_flare_304_aa_evolv.eps}
    \end{center}
    \caption{Sequence of AIA 304~{\AA}~images for the Nov 4th, 2015 flare. The white dotted box shows the portion of the IRIS raster FOV which scanned the ribbons. In panels b (c) F1 and F2 are the filament material, which move away from each other as the flare progresses. R1 and R2 in b (c,d) show the primary ribbons which move through the IRIS raster FOV. An animation of this image sequence is available in \cite{roy24}.}
    \label{flare_m_ev}
\end{figure*}
%%%--------------

Fig.\ref{flare_m_aia} presents the same region as depicted in Fig.\ref{flare_m_ev} across the six coronal channels (i.e., 94, 131, 171, 193, 211, 335~{\AA}) of SDO/AIA, recorded at the peak (top two rows) and during the decline phase (bottom two rows) of the flare. Post-eruption arcades \citep[][]{TriBC_2004} are clearly observable in all the channels, displaying slightly varied morphologies. These arcades contain evaporated thermal plasma and exhibit loop top brightening, likely due to colliding evaporation flows \citep[see, e.g.,][]{sharma16, patsourakos04}.

%%%--------------
\begin{figure*}[ht!]
    \begin{center}
        \includegraphics[trim={4.5cm 5.5cm 6cm 4cm},clip,width=0.9\textwidth]{Figures/nov_flare_aia_waves_2.pdf}
    \end{center}
    \caption{The Nov 4th, 2015 flare in the six Coronal channels of the SDO/AIA at the soft X-ray peak (panels (a)-(f)) and during the decline phase (panels (g)-(l)) of the flare.}
    \label{flare_m_aia}
\end{figure*}
%%%--------------

We begin by aligning the HMI observations with the IRIS observations using the 1600~{\AA} data recorded by AIA. Given that the observation involves an active region undergoing flaring, the magnetic field may also be rapidly changing \citep{wang02,dandan16,spirock02}. Therefore, we utilize a series of co-aligned full-disk maps from AIA and HMI to derive a rastered line-of-sight (LOS) map of magnetic flux density that precisely corresponds to the location and time of IRIS rasters. This procedure is outlined below.

Initially, we align the AIA 1600~{\AA} observation with the HMI observation closest in time to the IRIS SJI (Slit-Jaw Imager) and raster observations. Using the `aia\_prep' tool available in the \textit{sswidl} distribution, we process the level 1 images to perform image registration and align AIA and HMI observations. Subsequently, we calculate the shift between AIA 1600~{\AA} and the SJI 1400~{\AA}. Using these calculated offsets, we co-align the magnetograms with the IRIS SJI 1400~{\AA} observations, given that the AIA 1600~{\AA} observation was already aligned with the HMI observations. We typically observe an offset of approximately $\sim$ 1.5\arcsec~between HMI and IRIS. These co-aligned HMI maps are then utilized to derive the rastered magnetograms, which can be directly compared with IRIS raster observations.

For analyzing the properties of the  \ion{Mg}{2} lines, we fit a double Gaussian profile to both the k and h lines with a linear background symmetric about the line core. If excess emission is detected compared to the fitted background and line profile for wavelengths lower than 2792~{\AA} and in-between 2798~{\AA} to 2800~{\AA}, we infer the presence of  \ion{Mg}{2} triplets in emission. A single Gaussian profile is fitted to the excess emission. It's important to note that this approach may overlook the triplets unless the emission is sufficiently strong. However, this method serves our purpose since our primary objective is to characterize the  \ion{Mg}{2} k and h line profiles.

Additionally, we exclude any pixels showing saturation in either of the lines. The uncertainty of the observed intensities is measured in DN (Data Numbers) using the method outlined in \S 2:1 of \cite{kerr15} and subsequently applied in the fitting procedure. The uncertainties of the fitted Gaussian profiles are incorporated while integrating the profile to obtain the uncertainty of the line intensities.

%%%%-------------------------
\begin{figure}[h!]
\centering  
\includegraphics[trim={7cm 5cm 6cm 10cm},width=0.8\textwidth]{Figures/m_flare_iris.pdf}
\caption{Obtained images in  \ion{Mg}{2} h (panel a) and k (panel b) and corresponding co-aligned and artificially rastered HMI LOS magnetic field map (panel c) for the M3.7 flare observed on November 4th, 2015. The magenta and lime green contours on panel c show the contours of  \ion{Mg}{2}~h (panel a) and k (panel b) intensity.} \label{fig:aligned_raster}
\end{figure}
%%%%-------------------------

%%%%%%%%%%%
\begin{figure*}[ht!]
    \centering
    \includegraphics[trim={0cm 3cm 0.5cm 3cm},clip,width=\textwidth]{Figures/pix_fit.pdf}
    \caption{Fit of the  \ion{Mg}{2} window for the pixel marked in northern ribbon (southern ribbon) in Fig.~\ref{fig:aligned_raster} in top panel (bottom panel) from various times during the evolution of the flare. }
    \label{fig:pix_fit_ribbon}
\end{figure*}
%%%%%%%%%%%

%%%%-------------------------
\begin{figure}[ht!]
\centering  
\includegraphics[trim={1.5cm 4cm 0.5cm 4cm},clip,width=0.8\textwidth]{Figures/m_flare_iris_pt2.pdf}
\caption{The 1{--}8~{\AA} GOES light curve of the flare over-plotted with the time variation of  \ion{Mg}{2} k/h line intensity ratio for the northern (black) and southern (green) asterisks marked in panels a, b \& c.}
\label{fig:aligned_iris_ratio}
\end{figure}
%%%%-------------------------

In Figure~\ref{fig:aligned_raster}, we present intensity maps obtained in  \ion{Mg}{2} h \& k (panels a \& b) alongside the rastered line-of-sight (LOS) magnetic field map in panel (c). The dashed magenta and solid green contours on the magnetic field maps represent the intensity from panels a and b, respectively. To investigate the  \ion{Mg}{2} k to h intensity line ratios over time, we selected two pixels—one in the northern ribbon and another in the southern ribbon, as indicated by crosses. Fig.\ref{fig:pix_fit_ribbon} displays the spectra overlaid with fits for the two pixels taken in the northern ribbon (top panel) and southern ribbon (bottom panel) obtained at different times.

We illustrate the  \ion{Mg}{2} k to h line intensity ratio derived from the two locations in the northern (black) and southern ribbons (green) in Fig.\ref{fig:aligned_iris_ratio}. Additionally, we overlay the GOES light curve of the flare obtained in 1.0{--}8.0{\AA} (red solid line). Our observations indicate that the intensity ratios exhibit a slightly increasing trend, albeit within the margin of errors, during the early phase of the flare. The ratio demonstrates a distinct peak around the midpoint of the impulsive phase, lasting for a very brief duration of approximately 150 seconds, before beginning to decrease. It's worth emphasizing that the pixel-by-pixel fitting of the  \ion{Mg}{2} h \& k line profiles provided a sufficient signature of the time evolution of the k/h line intensity ratio during the flare.

%%%%%%%--------%%%%%%%%%%%
\subsection{Investigating the dependence on the local magnetic field in the  \ion{Mg}{2} lines}
%%%%%%%---------%%%%%%%%%%

To explore the time evolution of the intensity ratio and its potential relationship with the photospheric magnetic field, we binned the spectra into various magnetic field bins from different flaring pixels. We selected the flaring pixels based on an intensity threshold relative to the peak intensity observed in IRIS rasters. A double Gaussian fit with a constant background was performed in a smaller wavelength window for both the  \ion{Mg}{2} h and k lines separately. This step was essentially carried out to mitigate the effects of the background and potential contributions from other spectral features.

Fig.~\ref{fig:fit_pix_fov} displays the  \ion{Mg}{2} k line profiles obtained at different times and averaged over different magnetic field bins. Each plot denotes the observation time and the magnetic field strength over which the spectra were averaged and fitted with a double Gaussian. Line intensities were computed by integrating the fitted Gaussian.

In Fig.~\ref{fig:optical_depth_m}, we present the k to h intensity ratios obtained in various bins of magnetic flux density at different times corresponding to different flare phases. The curve corresponding to 13:07:05 UT (red) represents the pre-flare phase, where the ratio is approximately 1.2. The ratio shows an increase during the impulsive phase, i.e., the curves corresponding to 13:23:44 UT (blue dotted) and 13:32:04 UT (green dashed). The ratio measured at 13:41:11 UT (magenta dashed), which is closest to the UV peak of the flare, exhibits the largest value across all magnetic field bins. Subsequently, during the decline phase, the ratio measured at 13:57:23 UT (black solid) falls below the values observed during the pre-flare phase. Additionally, it's noteworthy that there are no significant differences among the k to h ratios obtained in different bins of magnetic flux density at all times (including pre-flare, impulsive, peak, and decline phases) of the flare.

%%%%-----------------------
\begin{figure*}[ht!]
    \begin{center}
    \includegraphics[trim={0cm 3cm 0cm 3cm},clip,width=\textwidth]{Figures/binned_fit_fov.pdf}
    \end{center}
    \caption{Fitted line profiles for the binned spectra for various magnetic field strengths from various times.}
    \label{fig:fit_pix_fov}
\end{figure*}
%%%%-----------------------

%%----------------------------------------------------------------------------
\begin{figure}[ht!]
    \centering
    \includegraphics[trim={8cm 1cm 2cm 0.2cm},clip,width=0.9\textwidth]{Figures/Flare-m-optical-depth-2.jpeg}
    \caption{ \ion{Mg}{2} k to h line intensity ratio for various magnetic flux density bins at various time steps during the evolution of the flare.}
    \label{fig:optical_depth_m}
\end{figure}
%%----------------------------------------------------------------------------

To explore further, we study the evolution of k to h intensity ratio as a function of time during the flare. In Fig.~\ref{fig:optical_dep_ev_m}, we plot the time evolution of the k to h line intensity ratio averaged within the bins of different magnetic flux densities as a function of time. For better visibility the 20.9{--}184.4 G and 348.9{--}513.3 G points are offset by 30s and -30s, respectively. We also plot the GOES 1{--}8~{\AA} light curve with the red solid line. These plots conspicuously reveal that the intensity ratios rise sharply during the impulsive phase, from $\sim$ 1.20 to $\sim$ 1.28 right before the soft X-ray flux peaks as seen from GOES, and decreases very rapidly thereafter, to lower than pre-flare values $\sim$ 1.12. The typical uncertainty value for the line intensity ratio is $\sim$ 0.02. We further note that the evolution of the  \ion{Mg}{2} k to h intensity ratio is remarkably similar across various strengths of magnetic flux densities.

%%%----------------------------------------------
\begin{figure*}[ht!]
    \centering
    \includegraphics[trim={3cm 3cm 2cm 4cm},clip,width=\textwidth]{Figures/Nov-11-2015-optical-dep-ev-5.pdf}
    \caption{Time evolution of the  \ion{Mg}{2} k to h line intensity ratio obtained from averaged spectrum over the corresponding magnetic flux bin as labeled for the Nov 4, 2015 flare. For better visibility, the 20.9{--}184.4 G and 348.9{--}513.3 G points are offset by 30s and -30s respectively. Over-plotted red solid line displays the 1{--}8~{\AA} GOES X-ray light curve.}
    \label{fig:optical_dep_ev_m}
\end{figure*}
%%%----------------------------------------------

%%%----------------------------------------------
\subsection{X1.6 Flare Observed on Oct 22, 2014}
%%%----------------------------------------------

An X~1.6 flare occurred on October 22, 2014, observed in AR 12192. The flare commenced at 14:02 UT and peaked at 14:28 UT as observed by GOES. Fig.\ref{flare2}(a) illustrates the GOES flux plot of the flare in the 0.5{--}4~{\AA} range (blue) and the 1.0{--}8.0~{\AA} range (red). Figs.\ref{flare2}(b) \& (c) display an AIA 1600{\AA} image and the line-of-sight (LOS) magnetic flux density map obtained from HMI, respectively. The overlaid boxes in panels (b) and (c) indicate the IRIS SJI (Slit-Jaw Imager) field of view (FOV). IRIS observed this flare with a large 8-step coarse raster covering a FOV of [14\arcsec,174\arcsec]. It's worth noting that the IRIS SJI FOV and the slit direction were rotated by approximately 45$^\circ$ relative to the center of the HMI observation.

%%--------------------------------------------------
\begin{figure*}[ht!]
    \centering
\hspace*{-.5in}
\includegraphics[width=\textwidth,trim={7cm 5cm 7.7cm 6cm},clip]{Figures/Flare_X_oct22_2014_2.eps}
\caption{The X class flare observed on October 22, 2014. Panel a: GOES flux plot in 0.5{--}4~{\AA} (blue) and 1.0{--}8.0~{\AA} (red). Panel b: AIA 1600~{\AA} image taken at the peak of the flare. Arrows locate the primary and secondary ribbons. Panel c: LOS magnetic flux density map obtained from HMI at the peak of the flare. The over-plotted white (black) box in panel b (c) represents the IRIS SJI FOV, and white dot-dashed (dashed) box in panels b (c) represents the IRIS Raster FOV.}\label{flare2}
\end{figure*}
%%--------------------------------------------------

The flare manifests two ribbons in AIA 1600{\AA}, as indicated by the arrows. The eastern ribbon extends around a negative field region and is entirely covered by the IRIS SJI FOV. Conversely, the western ribbon extends around a positive field region, with only a portion of it covered by the IRIS SJI FOV, as depicted in panel (b). In Fig.~\ref{fig:align_raster_flare2}, we present intensity maps obtained in  \ion{Mg}{2} h \& k (panels a \& b) alongside the rastered line-of-sight (LOS) magnetic field map (panel c). The black \& green dashed contours in the LOS magnetic field map (panel c) indicate the intensity contours of  \ion{Mg}{2} h \& k (panels a \& b).
